### 2015-04-15 paper/ms.tex
paper template

%Document Class
\documentclass{latex/emulateapj}
\usepackage{multirow,color,wrapfig,ulem}

%Packages and Commands
\include{latex/packages_commands}

\begin{document}

\title{Titulo}
\shorttitle{shorttitle}

\shortauthors{shortauthors}

\author{author1, author2 \& author3} 
\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia}

\email{email1}
\email{email2}
\email{email3}

\keywords{keywords....}

\begin{abstract}

Abstract...

\end{abstract}

\section{Introduction}
\label{sec:intro}

Introduction... \\

\section{Theoretical Background}
\label{sec:theo}

Theoretical background...\\

\section{Results}
\label{sec:results}

Results...\\

\section{Discussion}
\label{sec:discussion}

Discussion...\\

\section{Conclusions}
\label{sec:conclusions}

Conclusions...\\

\section*{Acknowledgments}

Acknowledgments...\\

\bibliographystyle{latex/apj}
\bibliography{references}

\newpage