### 2015-04-29 paper/ms.tex

\documentclass{emulateapj}
\submitted{{\it Submitted for publication in ApJ Letters}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}
\usepackage{graphics}
\usepackage[dvips]{epsfig}
\usepackage{float}
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}   
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}   
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman $\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}
% definition to produce a "less than or similar to" symbol
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}
% definition to produce a "greater than or similar to" symbol
\def\gsim{~\rlap{$>$}{\lower 1.0ex\hbox{$\sim$}}}
%@arxiver{fig3.pdf,fig11a.pdf, fig11b.pdf} 
\begin{document}

\title{First estimate of a galaxy rotational velocity from
  its observed Lyman-$\alpha$ line} 
\shorttitle{Rotational velocity from the Lyman $\alpha$ line}

\shortauthors{Remolina-Gutierrez, Garavito-Camargo, Forero-Romero}

\author{ Maria C. Remolina-Gutierrez, Juan N. Garavito-Camargo, Jaime
  E. Forero-Romero}  
\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
\email{mc.remolina197@uniandes.edu.co}
\email{jn.garavito57@uniandes.edu.co}
\email{je.forero@uniandes.edu.co}

\keywords{galaxies: high-redshift --- line: formation --- methods:
  numerical ---  radiative transfer} 
\begin{abstract}
First.
\end{abstract}

\section{Introduction}
\label{sec:intro}

Introduction... \\ 
\cite{Garavito14}

\subsection{Theoretical Background}
\label{sec:theo}

\cite{Garavito14} presented an analytical expresion that approximates
the results for the Lyman-$\alpha$ line morphology for a spherical galaxy
rotating as a solid body.

Ths input parameters of the analytical solution are the optical depth
$\tau_{H}$, the rotational velocity at the sphere's surface $V_{\rm
  max}$ and the angle of view measured from the equatorial plane
(which is perpendicular to the rotational axis) $\theta$. 
A $\texttt{C}$ source code implementation is available in the public
repository associated to this paper.

The main result \cite{Garavito14} is that rotation does have a
noticeable impact on the \ly line morphology.
This impact is two fold. 
First, the rotation axis breaks the spherical symmetry, introducing a
dependence of the line shape with viewing angle.
Second, for increasing rotational velocities and viewing angles off
the rotational axis, the intensity at the line's centers increases;
for sufficiently high rotational velocities the line becomes single peaked.

An approximated analytic solution was derived for a spherical galaxy with 
central \lya sources.

\begin{equation}
J(x,b,\phi,i)=\frac{\sqrt{\pi}}{\sqrt{24}a\tau_0}\Bigg{(}\frac{(x-x_{\rm
    b})^2}{1+{\rm cosh}\Big{[}\sqrt{\frac{2\pi^3}{27}}\frac{|(x
      -x_{\rm b})^3|}{a\tau_0}\Big{]}}\Bigg{)},
\end{equation}

\begin{equation}
J(x,i) \approx 2\pi \int_0^Rdb \hs b
\int_0^{2\pi}d\phi \hs J(x,b,\phi,i)
\end{equation}

\subsection{Low redshift observations}

The \ly emission line is a powerfull tool to study the properties 
of the high redshift Universe. Important intrinsic properties of this
galaxy population, such as the redshift, luminosity functions, 
equivalent width and the kinematic state are derived from the 
\ly line. With the aim to study in more detail the impact of this 
properties on the \ly line several surveys have observe nearby
galaxies whose properties are known from . We are particularly
interested in study the kinematic state of the galaxys using 
the \ly line. The \citep{LARS} survey have observed 14 nearby galaxies
with strong \ly emission using the HST telescope and the .

The LARS survey have observed galaxies with different properties, 
the spectra of all this galaxies and their properties are described 
in detail in \citep{LARS2} \citep{LARS3, LARS4}.

\section{Results}
\label{sec:results}

$$\tau_{H} = 69663.485 $$
$$V_{\rm max} = 49.961\kms $$
$$\theta = 44.495^\circ $$
$$T = 9962.731 \rm K$$
$$\Delta V_{off} = -9.722 \kms$$

\begin{figure}[H]
\begin{center}
  \includegraphics[width=0.5\textwidth]{mcmc.png}
\end{center}
\caption{
    \label{fig:mcmc_result}}  
\end{figure}

\begin{figure}[H]
\begin{center}
  \includegraphics[width=0.5\textwidth]{parameters.png}
\end{center}
\caption{
    \label{fig:mcmc_triangle}}  
\end{figure}

Results...\\

\section{Discussion}
\label{sec:discussion}

Discussion...\\

\section{Conclusions}
\label{sec:conclusions}

Conclusions...\\

\section*{Acknowledgments}

Acknowledgments...\\

\bibliographystyle{apj}
\bibliography{references}

\end{document}